What We Eval in the Shadows

### 2021-08-11 paper/main.tex
Fix
Removed example from min

\begin{tabular}{@{}p{.97\linewidth}}
  \vspace{1mm}
  \EE{$min(e)=\code{V}$} \\[-2mm]\small Values occur in 16\% of sites, 74\% of
  those are constants such as integers, the remainder evaluate to a value. \Eval
  is needed when the expression denoting the values is constructed
  programmatically. Cases when a value is directly given to \eval, and \eval
  returns it unchanged, often occur when a computation has a default path and
  another, more interesting, path that requires evaluation. The following is a
  call site that always see a string; \packageNbCallSitesUniqueActualValue of
  sites only ever see a value. This shape usually run in few interpreter steps.
  The environment in which they evaluate is mostly irrelevant.
\begin{lstlisting}
  eval(paste("A~",paste(paste(names(X[,-1])),collapse="+")))
\end{lstlisting}
\vspace{2mm}
\EE{$min(e)=\code{X}$}\\[-2mm]\small Variable lookups are the most common shape,
\packageNbSymbolVarSitePercent are simple variable names, e.g. \c x. The shapes
subsumes \c{V} and built-in arithmetic operations, so a mixture of arithmetics
and variables normalize to \c X. A lookup is one step of execution, if a promise
is returned an arbitrary number of additional steps may be needed; the median is
\packageMinimizedmedianoperationsaRnd, suggesting that most variable read do
little work. Variables are often evaluated in environments that have been
constructed programmatically, \packageMinimizedpercentparentframesa of
expressions are evaluated in a function environment.
\vspace{2mm}
\EE{$min(e)=\code{\$}$}\\[-2mm]\small This shape extends \c X to include lookup
with the dollar operator, \eg~\c{x\$f}, and vector indexing, \eg~\c{x[42]} and
\c{x[[24]]}. This takes a few more step of evaluation on average, and is
typically used in a function environment.

\vspace{2mm}
\EE{$min(e)=$~\code{<-}}\\[-2mm]\small This includes both direct
assignment {\tt <-} and parent environment assignment {\tt <\,\!<-} as well as
the \c{assign} function. It subsumes \c{\$}. Assignments represent the most
obvious source of side effects in \eval. Most cases originate from trivial code
generation where the assign expression is assembled using parse or substitute,
often in a loop. For example (taken from \c{PopED::get\_rse}), the following
code will populate the current scope with a number of variables:
%
\begin{lstlisting}
for (i in seq_along(args)) eval(parse(text=paste(args[i], "<-", i)))
\end{lstlisting}

\vspace{2mm}
\EE{$min(e)=\code{F()}$}\\[-2mm]\small Function calls without variable lookups or
assignment. A variable is allowed the function position, looking up
functions does not trigger computation unless the function's name
is bound to a promise. This shape typically does not perform much work.
\vspace{2mm}
\framebox{$min(e)=\code{F(X)}$}~\EE{$min(e)=\code{F(F(X))}$}\\[-2mm]\small Function
calls whose arguments may include variable references and nested calls. Together
they represent the most frequent shape and perform over 16 steps. They are
frequently evaluated in function environments.
\end{tabular}

@@

Out of the minimized expressions, it is \c{BLOCK}, \c{<-} and \c{F(F(X)} that
contribute to the vast majority of side effects. Concretely, 92\% of \c{BLOCK},
87\% of \c{<-} and 15\% of \c{F(F(X))} expressions do side effects. In the case
of \c{BLOCK}, most of them happen in the environment passed to the \eval call
(default is the parent environment). For \c{<-} it is 15\%. However, this form
is dominated by the \c{plyr::allocate\_column} function that side-effects in
the environment containing the data frame to which it allocates new column.
Without this call, \c{<-} is similarly predictable with 96\% of side effects
happening in the passed environment. For \c{F(F(X))}, it is less clear as 58\%
of side effects happen in a different environment to the one that was passed to
\eval.

>>>>

\begin{tabular}{@{}p{.97\linewidth}}
  \vspace{1mm}
  \EE{$min(e)=\code{V}$} \\[-2mm]\small Values occur in 16\% of sites, 74\% of
  those are constants such as integers, the remainder evaluate to a value (\packageNbCallSitesUniqueActualValue of
  sites only ever see a value). \Eval
  is needed when the expression denoting the values is constructed
  programmatically. Cases when a value is directly given to \eval, and \eval
  returns it unchanged, often occur when a computation has a default path and
  another, more interesting, path that requires evaluation.
  This shape usually run in few interpreter steps.
  The environment in which they evaluate is mostly irrelevant.

  \vspace{2mm}
  \EE{$min(e)=\code{X}$}\\[-2mm]\small Variable lookups are the most common shape,
  \packageNbSymbolVarSitePercent are simple variable names, e.g. \c x. The shapes
  subsumes \c{V} and built-in arithmetic operations, so a mixture of arithmetics
  and variables normalize to \c X. A lookup is one step of execution, if a promise
  is returned an arbitrary number of additional steps may be needed; the median is
  \packageMinimizedmedianoperationsaRnd, suggesting that most variable read do
  little work. Variables are often evaluated in environments that have been
  constructed programmatically, \packageMinimizedpercentparentframesa of
  expressions are evaluated in a function environment.
  \vspace{2mm}
  \EE{$min(e)=\code{\$}$}\\[-2mm]\small This shape extends \c X to include lookup
  with the dollar operator, \eg~\c{x\$f}, and vector indexing, \eg~\c{x[42]} and
  \c{x[[24]]}. This takes a few more step of evaluation on average, and is
  typically used in a function environment.

\vspace{2mm}
\EE{$min(e)=$~\code{<-}}\\[-2mm]\small This includes both direct
assignment {\tt <-} and parent environment assignment {\tt <\,\!<-} as well as
the \c{assign} function. It subsumes \c{\$}. Assignments represent the most
obvious source of side effects in \eval. Most cases originate from trivial code
generation where the assign expression is assembled using parse or substitute,
often in a loop.

\vspace{2mm}
\EE{$min(e)=\code{F()}$}\\[-2mm]\small Function calls without variable lookups or
assignment. A variable is allowed the function position, looking up
functions does not trigger computation unless the function's name
is bound to a promise. This shape typically does not perform much work.

\vspace{2mm}
\framebox{$min(e)=\code{F(X)}$}~\EE{$min(e)=\code{F(F(X))}$}\\[-2mm]\small Function
calls whose arguments may include variable references and nested calls. Together
they represent the most frequent shape and perform over 16 steps. They are
frequently evaluated in function environments.
\end{tabular}

@@

Out of the minimized expressions, it is \c{BLOCK}, \c{<-} and \c{F(F(X))} that
contribute to the vast majority of side effects. Concretely, 92\% of \c{BLOCK},
87\% of \c{<-} and 15\% of \c{F(F(X))} expressions do side effects. In the case
of \c{BLOCK}, most of them happen in the environment passed to the \eval call
(default is the parent environment). For \c{<-} it is 15\%. However, this form
is dominated by the \c{plyr::allocate\_column} function that side-effects in
the environment containing the data frame to which it allocates new column.
Without this call, \c{<-} is similarly predictable with 96\% of side effects
happening in the passed environment. For \c{F(F(X))}, it is less clear as 58\%
of side effects happen in a different environment to the one that was passed to
\eval.
